Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-09 arxiv/proof_examples.tex
B.2 polish

We provide the proof of Corollary \ref{cor_MTL_loss}. The proof is similar to that of Theorem \ref{thm_many_tasks}.
First, we reduce the optimization objective to another function $g(x)$ that only depends on the sample covariance matrices.
Second, we show that the global minimizer of $g(x)$ is approximately equal to $1$.
Finally, we use the above result and Theorem \ref{thm_main_RMT} to prove the proposition.

In the setting of Corollary \ref{cor_MTL_loss}, the optimization objective \eqref{eq_mtl} is equal to

@@

Let denote $\hat \Sigma(x):= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}$.

@@

Using this identity and taking expectation over $\epsilon^{(1)}$ and $\epsilon^{(2)}$, we can simplify \eqref{Eg(x)}  to

@@

For the random-effects model, we take the expectation over $\beta_1$ and $\beta_2$:

@@

where we used that the entries of $\beta_1-x\beta_2\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $p^{-1}[(x-1)^2\kappa^2 + (x^2+1)d^2/2]$.

\paragraph{Part $1$: approximating the global minimum.}
Let $\hat{x}$ denote the minimizer of $g(x)$.
We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.

First, using the concentration estimates in Lemma \ref{largedeviation}, we show that $g(x)$ and its conditional expectation over $X^{(1)}, X^{(2)}$ are close. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(x) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability
$$\left| g(x) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(x) \mid X_1, X_2 }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right)$$.
%for any small constant $c>0$.
	% \HZ{use consistent $e_1, e_2, e$, etc. Why not just use $\e$? Don't create too many different symbols. It's bad style..}
 \end{claim}
 \begin{proof}
By Fact \ref{lem_minv} (iii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
 \ee
 and
 \be\label{op_Sig1}
\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
 \ee
where we used that $1+\tau\le \rho_1=n_1/p\le \tau^{-1}$ and $1+\tau\le \rho_2=n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

@@

Finally, we can obtain similar estimates for the second term on the right-hand side of \eqref{eq_mtl_A12} as in  \eqref{eq_hA1}, \eqref{eq_hA23}, \eqref{eq_hA4} and \eqref{eq_hA56}, which conclude the proof of Claim \ref{claim_largedev1}.
 \end{proof}

Next, bu studying the conditional expectation in \eqref{EgA}, we show that $\hat x$ is close to 1.

\begin{claim}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability,
%There exists a constant $C>0$ such that
	\be\label{hatw_add1}|\hat x -1|\le  \frac{2d^2}{\kappa^2} + p^{-1/4+c}.
	\ee
%	for any small constant $c>0$.
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{claim}

@@

\paragraph{Part $2$: reducing the random-effects model to Theorem \ref{thm_main_RMT}.}
%\HZ{restructure this part. merge the result of B.17 and B.18, and state it upfront. separate its proof into two parts}
%\HZ{just say that this calculation of B.97 is similar to B.81?}
We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat x=\hat x_1/\hat x_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|\Sigma_2^{1/2} \left( \hat B \hat x_2 - \beta_2\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
%where we denoted $$\hat \Sigma:= \hat x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
Using Lemma \ref{lem_hat_v} and the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as in the following claim.

\begin{claim}
In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
%\be\label{choiceofpara}
%	\frac{pd^2}{\sigma^2} = \OO(1),\quad p^{-1+c_0} \le \frac{\kappa^2}{\sigma^2 }  \le p^{-\e_0-c_0},
%\ee
%where $\e_0$ is the constant as appeared in equation \eqref{approxvalid}. Then we have
\begin{align}
&\left|L(\hat{\beta}_2^{\MTL}) - d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat \Sigma^{-1}  }\right)\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) , \label{simple1}
\end{align}
for any small constant $c>0$, where we recall that $\hat \Sigma=\hat \Sigma(1)$ was defined in \eqref{def hatsig}.
\end{claim}
\begin{proof}
 \iffalse
 To illustrate the idea of bias-variance decomposition, we assume that $\hat x$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2}  &=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(2)})^\top} \nonumber\\
			&+ \sigma^2\cdot \hat x^2 \bigtr{ X^{(1)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 +\sigma^2  \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	& \exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2} \nonumber\\
	&	=  \hat x^2 \left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2 \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma^{(2)} (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

%Plugging $Y^{(1)}$ and $Y^{(2)}$ in equation \eqref{model_YvsX} into equation \eqref{eq_Bhat_2task}, we can rewrite that 
% \begin{align}
%		\hat{B}= \frac{1}{A_2}\left[\beta_2 + \left(\frac{A_1^2}{A_2^2}(X^{(1)})^{\top}X^{(1)} +  (X^{(2)})^{\top}X^{(2)}\right)^{-1}\bigbrace{ (X^{(1)})^{\top}X_1\left(\frac{A_1}{A_2}\beta_1 - \frac{A_1^2}{A_2^2} \beta_2\right) + \left(\frac{A_1}{A_2} (X^{(1)})^{\top}\varepsilon_1 + (X^{(2)})^{\top}\varepsilon_2\right)}\right]. \label{eq_Bhat_2task2} %\\
%		%&= (B^\star A ^{\top}) (A A^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top}   \bigbrace{\sum_{j=1}^t \varepsilon_i A_i^{\top}} (A  A^{\top})^{-1}.
%	\end{align}

In general, for $\hat x$ as  the minimizer of $g(x)$, it depends on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. However, using the concentration estimates in Lemma \ref{largedeviation}, we can still obtain the expression in equation \eqref{eq_var_2task} up to a small error. In particular, our proof will not use Lemma \ref{lem_hat_v}.
\fi
Our proof is divided into two steps. First, with the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as follows. With high probability, we have that for any small constant $c>0$,
%control the difference between $g(x)$ and its partial expectation. 
% \begin{claim}\label{claim_largedev2}
 %In the setting of Proposition \ref{cor_MTL_loss}, if $\hat x=\OO(1)$, then we have that 
\be\label{claim_largedev2} \left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat x)\right| \le p^{-1/2+c} \left(\sigma^2 +\kappa^2 +d^2 \right),
\ee
where the function $\cal L(\hat x)$ is defined as
\begin{align*}
	\cal L(\hat x)	:=  &\hat x^2\left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} } \\
	& +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
\end{align*}
% \end{claim}
We postpone the proof of equation \eqref{claim_largedev2} until we complete the proof of \eqref{simple1}. Next, we can further simplify $\cal L(\hat x)$ using Claim \ref{lem_hat_v} and $\Sigma^{(1)}=\Sigma^{(2)}=\id_p$.
More precisely, we claim that with high probability,
%\begin{claim}
%In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
%\be\label{choiceofpara}
%	\frac{pd^2}{\sigma^2} = \OO(1),\quad p^{-1+c_0} \le \frac{\kappa^2}{\sigma^2 }  \le p^{-\e_0-c_0},
%\ee
%where $\e_0$ is the constant as appeared in equation \eqref{approxvalid}. Then we have
\begin{align}
&\left|\cal L(\hat x)- d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat \Sigma^{-1}  }\right)\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) , \label{simple12}
\end{align}
for any small constant $c>0$. Combining \eqref{claim_largedev2} and \eqref{simple12}, we immediately obtain \eqref{simple1}. %, where we recall that $\hat \Sigma=\hat \Sigma(1)$ as defined in \eqref{def hatsig}.
%\end{claim}
%\begin{proof}

It remains to prove \eqref{simple12}. Using Lemma \ref{lem_hat_v} and equations \eqref{op_X12}-\eqref{op_Sig1}, we obtain that with high probability,
 \|\hat\Sigma^{-1}-\hat\Sigma(\hat x)^{-1}\| &\le |\hat x^2-1|\|\hat\Sigma^{-1}\| \| (X^{(1)})^\top X^{(1)}\|\|\hat\Sigma(\hat x)^{-1}\|  \lesssim p^{-1}\left(\frac{d^2}{\kappa^2} + p^{-1/4+c}\right),\label{est111}
for any small constant $c>0$. With similar arguments, we can get that with high probability,
\begin{align}\label{est222}
&\left\|\left(\hat\Sigma^{-2}-\hat\Sigma(\hat x)^{-2}\right)\left((X^{(1)})^\top X^{(1)} \right)^2\right\| \lesssim  \frac{d^2}{\kappa^2} + p^{-1/4+c} .
Moreover, using equations \eqref{op_X12}-\eqref{op_Sig1} we can bound that with high probability,
%$$\bignorm{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)}}\lesssim 1.$$
%which further implies that with high probability,
& \bigtr{ \hat\Sigma^{-2} \left((X^{(1)})^{\top}X^{(1)}\right)^2}  \le p \bignorm{\hat\Sigma^{-2} \left((X^{(1)})^{\top}X^{(1)}\right)^2}\lesssim p.\label{est333}
Now we bound the left-hand side of \eqref{simple1} as % \HZ{replace $\Sigma(1)$ with $\hat{\Sigma}$ throughout}
&\left|\cal L(\hat x)-\frac{d^2}{p}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat\Sigma^{-1}  }\right)\right| \\
&\lesssim \left(|\hat x-1|^2 \kappa^2 + |\hat x-1|d^2\right)\cdot p^{-1}\bigtr{ \hat\Sigma^{-2} \left((X^{(1)})^{\top}X^{(1)}\right)^2 } \\
&+ \frac{d^2}{p}\left|\tr\left[\left(\hat\Sigma(\hat x)^{-2}-\hat\Sigma^{-2}\right) \left((X^{(1)})^\top X^{(1)} \right)^2\right]\right|+ \sigma^2  \left|\bigtr{\hat\Sigma(\hat x)^{-1}-\hat\Sigma^{-1}  }\right|.
Applying the estimates \eqref{hatw_add1}, \eqref{est111}, \eqref{est222} and \eqref{est333} to the three terms on the right-hand side, we can conclude \eqref{simple12}.
% Inserting equation \eqref{hatv_add1} into equation \eqref{eq_te_mtl_2task} and applying Lemma  \ref{largedeviation} to $(\beta_1-\hat v\beta_s)$, we get that w.h.p.,

>>>>

We provide the proof of Corollary \ref{cor_MTL_loss}.
We characterize the global minimizer of $f(A, B)$ in the random-effects model.
Based on the characterization, we reduce the prediction loss of hard parameter sharing to the asymptotic limits provided in Theorem \ref{thm_main_RMT}.
Then, we prove Corollary \ref{cor_MTL_loss} based on the bias and variance limits.
We set up several notations.
In the two-task case, the optimization objective $f(A, B)$ is equal to

@@

We denote $\hat \Sigma(x):= x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}$.

@@

Using this identity, we can simplify equation \eqref{Eg(x)}  to

@@

For the random-effects model, recall that the entries of $\beta_1, \beta_2\in \R^p$ are i.i.d. Gaussian random variables with mean zero and variance $\kappa^2/p, d^2/p$, respectively.
Hence,

@@

\paragraph{Part $1$: characterizing the global minimum of $f(A, B)$.}
Let $\hat{x}$ denote the minimizer of $g(x)$.
We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.
This gives us the global minimum of $f(A, B)$, combined with the local optimality condition for $\hat{B}$.

First, we show that $g(x)$ and $\ex{g(x) \mid X^{(1)}, X^{(2)}}$ are close using standard concentration bounds.
\begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability
$$\left| g(x) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(x) \mid X_1, X_2 }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right).$$
%for any small constant $c>0$.
 \end{claim}
 \begin{proof}
By Fact \ref{lem_minv} (iii), we have that with high probability,
 \be\label{op_X12}
\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
 \ee
 and
 \be\label{op_Sig1}
\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
 \ee
where we used that $1+\tau\le \rho_1=n_1/p\le \tau^{-1}$ and $1+\tau\le \rho_2=n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

@@

Finally, we can obtain similar estimates for the second term on the right-hand side of \eqref{eq_mtl_A12} as in  \eqref{eq_hA1}, \eqref{eq_hA23}, \eqref{eq_hA4} and \eqref{eq_hA56}. Hence the proof is complete.
 \end{proof}

Next, we show that $\hat x$ is close to 1.
\begin{claim}\label{lem_hat_v}
%Suppose the assumptions of Lemma \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order.
 In the setting of Corollary \ref{cor_MTL_loss}, we have that with high probability,
%There exists a constant $C>0$ such that
	\be\label{hatw_add1}|\hat x -1|\le  \frac{2d^2}{\kappa^2} + p^{-1/4+c}.
	\ee
%	for any small constant $c>0$.
	%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{claim}

@@

\paragraph{Part $2$: a reduction to the bias and variance limits.}
Recall that the hard parameter sharing estimator $\hat{\beta}_2^{\MTL}$ is equal to $\hat{B} \hat{A}_2$.
The predication loss of hard parameter sharing is as follows
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|\Sigma_2^{1/2} \left( \hat B \hat A_2 - \beta_2\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
Using Lemma \ref{lem_hat_v} and the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as in the following claim.

\begin{claim}\label{claim_reduction}
Recall that $\hat \Sigma(1)$ is equal to $\hat \Sigma$ (cf. Section \ref{sec_diff}).
In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability
\begin{align*}
&\left|L(\hat{\beta}_2^{\MTL}) - d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \bigtr{{\hat \Sigma^{-1}  }}\right| \nonumber\\
\lesssim&  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2).
\end{align*}
\end{claim}
\begin{proof}
Our proof is divided into two steps. First, with the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as follows. With high probability, we have that for any small constant $c>0$,
\be\label{claim_largedev2} \left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat x)\right| \le p^{-1/2+c} \left(\sigma^2 +\kappa^2 +d^2 \right),
\ee
where the function $\cal L(\hat x)$ is defined as
\begin{align*}
	\cal L(\hat x)	:=  &\hat x^2\left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} } \\
	& +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
\end{align*}
% \end{claim}
We are ready to finish the proof of Claim \ref{claim_reduction}.
We defer the proof of equation \eqref{claim_largedev2} until the end.
Next, we can further simplify $\cal L(\hat x)$ using Claim \ref{lem_hat_v} and $\Sigma^{(1)}=\Sigma^{(2)}=\id_p$.
More precisely, we claim that with high probability
\begin{align}
&\left|\cal L(\hat x)- d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \bigtr{\hat \Sigma^{-1}  }\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2). \label{simple12}
\end{align}
Combining \eqref{claim_largedev2} and \eqref{simple12}, we obtain Claim \ref{claim_reduction}.

It remains to prove equation \eqref{simple12}. Using Lemma \ref{lem_hat_v} and equations \eqref{op_X12}-\eqref{op_Sig1}, we obtain that with high probability,
\begin{align}
 \|\hat\Sigma^{-1}-\hat\Sigma(\hat x)^{-1}\| &\le |\hat x^2-1|\|\hat\Sigma^{-1}\| \| (X^{(1)})^\top X^{(1)}\|\|\hat\Sigma(\hat x)^{-1}\|  \lesssim p^{-1}\left(\frac{d^2}{\kappa^2} + p^{-1/4+c}\right).\label{est111}
\end{align}
Using similar arguments, we get that with high probability,
\begin{align}\label{est222}
&\left\|\left(\hat\Sigma^{-2}-\hat\Sigma(\hat x)^{-2}\right)\left((X^{(1)})^\top X^{(1)} \right)^2\right\| \lesssim  \frac{d^2}{\kappa^2} + p^{-1/4+c} .
\end{align}
Moreover, using equations \eqref{op_X12}-\eqref{op_Sig1} we can bound that with high probability,
%$$\bignorm{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)}}\lesssim 1.$$
%which further implies that with high probability,
\begin{align}
& \bigtr{ \hat\Sigma^{-2} \left((X^{(1)})^{\top}X^{(1)}\right)^2}  \le p \bignorm{\hat\Sigma^{-2} \left((X^{(1)})^{\top}X^{(1)}\right)^2}\lesssim p.\label{est333}
\end{align}
Now we bound the left hand side of equation \eqref{simple12} as
\begin{align*}
&\left|\cal L(\hat x)-\frac{d^2}{p}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat\Sigma^{-1}  }\right)\right| \\
&\lesssim \left(|\hat x-1|^2 \kappa^2 + |\hat x-1|d^2\right)\cdot p^{-1}\bigtr{ \hat\Sigma^{-2} \left((X^{(1)})^{\top}X^{(1)}\right)^2 } \\
&+ \frac{d^2}{p}\left|\tr\left[\left(\hat\Sigma(\hat x)^{-2}-\hat\Sigma^{-2}\right) \left((X^{(1)})^\top X^{(1)} \right)^2\right]\right|+ \sigma^2  \left|\bigtr{\hat\Sigma(\hat x)^{-1}-\hat\Sigma^{-1}  }\right|.
\end{align*}
Applying the estimates \eqref{hatw_add1}, \eqref{est111}, \eqref{est222} and \eqref{est333} to the three terms on the right-hand side, we can conclude \eqref{simple12}.

Now we give the proof of the estimate \eqref{claim_largedev2}.
The proof of this claim is very similar to Claim \ref{claim_largedev1}, where the only difference is that $\hat x$ now may depend on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$.
In this case, we can still use  Lemma \ref{largedeviation} to conclude the proof because $\hat x$ can be pulled out as a coefficient.
Here we only give a proof sketch and omit the details.
Recall that $\beta_0$ is the shared component of $\beta_1$ and $\beta_2$ with i.i.d. Gaussian entries of mean zero and variance $p^{-1}\kappa^2$. Moreover, we denote the task-specific components by $\wt\beta_1$ and $\wt\beta_2$, whose entries are i.i.d. Gaussian random variables of mean zero and variance $p^{-1} d^2 / 2$. Then we write $L(\hat{\beta}_2^{\MTL}) $ in \eqref{Lbeta_HPS} as:
\begin{align}
L(\hat{\beta}_2^{\MTL})  =&  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x -\hat x^2)\beta_0+(X^{(1)})^\top X^{(1)} \hat x\wt \beta_1 - (X^{(1)})^\top X^{(1)}  \hat x^2\wt \beta_2 \right] \right. \nonumber\\
&\left. + (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1}\left[ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2. \label{expand_15}
\end{align}
As in \eqref{expand_6}, we can expand this expression into the sum of 15 terms, and bound each term as in \eqref{eq_hA1}-\eqref{eq_hA56}. For example, for the main term $\hat x^2 \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1$, using 
Lemma \ref{largedeviation} and equation \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability:
\begin{align*}
&  \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1 - \frac{d^2}{2} \cdot p^{-1}\bigtr{(X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)}}\right| \\
&\le p^{-1+c}d^2 \cdot \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)} \right\|_F \\
&\le p^{-1/2+c}d^2 \cdot  \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)} \right\|  \lesssim p^{-1/2+c}d^2.
\end{align*} 
For the cross term $\hat x \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \epsilon^{(2)}$, using Lemma \ref{largedeviation} and \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability for any small constant $c>0$:
\begin{align*}
 \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \epsilon^{(2)}\right|  & \le p^c \cdot \sigma\sqrt{p^{-1}d^2} \cdot \|(X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \|_F \\
&\lesssim p^{c}\sigma d \cdot  \|(X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \| \\
&\lesssim p^{-1/2+c}\sigma d\le p^{-1/2+c}(\sigma^2+ d^2).
\end{align*}
The rest of the terms in the expansion of \eqref{expand_15} can be bounded in the same way, and we omit the details.
 \end{proof}

\paragraph{Part $3$: applying the bias and variance limits.}
Finally, we are ready to complete the proof of Corollary \ref{cor_MTL_loss}.
We derive the variance term $\sigma^2  \tr[\hat\Sigma^{-1}]$ and the bias term $d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right]$ using Theorem \ref{thm_main_RMT}.

\begin{proof}[Proof of Corollary \ref{cor_MTL_loss}]
Using \eqref{lem_cov_shift_eq}, we obtain that
\be\label{eq_var111}\tr[\hat\Sigma^{-1}] = \tr\left[ \left((X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}\right)^{-1}\right]=\bigtr{\frac{(a_1\id_p+a_2\id_p)^{-1}}{n_{1}+n_2} }+\OO(p^{-c_\varphi})\ee
with high probability. Solving equation \eqref{eq_a12extra} with $\lambda_i\equiv 1$, $1\le i\le p$, we get that  
	\begin{align}
		 a_1 = \frac{\rho_1(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(\rho_1 + \rho_2 - 1)}{(\rho_1 + \rho_2)^2} . \label{simplesovlea12}
			\end{align}
Applying the above to equation \eqref{eq_var111}, we obtain that
\be\label{eq_var112}\bigtr{\hat\Sigma^{-1}}  = \frac{p}{n_1+n_2} \cdot \frac{\rho_1+\rho_2}{\rho_1+\rho_2-1}+\OO(p^{-c_\varphi})=  \frac{1}{\rho_1+\rho_2-1}+\OO(p^{-c_\varphi})\ee
with high probability.
